---
title: "Preparing and labelling probe from a complex oligo library"
author: "omidalam"
date: "Last updated: 8/8/2018"
output: html_document
---

```{r setup, echo =F}
library(knitr)
library(kableExtra)
```

---

## Introduction:

This protocol employs emulsion PCR (ePCR) to enrich sub libraries from a complex oligo library. I have prepared this protocol based on [ _"Amplification of complex gene libraries by emulsion PCR"_ ](https://www.ncbi.nlm.nih.gov/pubmed/16791213) _R. William, et.al, Nature Methods_[^1] and _"Preparing Single-Stranded Labelled Probes from MYtags Immortal Lirary"_[^2], Arbor Biosciences.
Labelling protocol is based terminal transferase protocol from Roche. 

### Materials
+ Oligo nucleotide library
+ Enrichment primer sets
+ Nuclease-free water
+ Span 80 (Sigma S6760)
+ Tween 80 (Sigma P5188)
+ Triton X-100 (Sigma T8787)
+ Minral Oil (Sigma M5904)
+ Bovine Serum Albumin (BSA) (Sigma B6917)
+ diethyl ether 


### Equipments
+ BRAND magnetic stirring bar, PTFE, cylindrical with pivot ring (Sigma Z329061). pre-cleaned (See ["Cleaning stir bars"](./cleaning_stir_bars.html)).
+ Magnetic stirrer with speed control (Burrell Scientific)
+ PCR cleanup kit

-----

###Before you begin!

#### Prepare library
1. Prepare a stock solution of your library. 
2. Resuspend at 1 ng/μl in 10 mM Tris-HCl pH 7.5 or nuclease-free water. 
    + Keep the stock at -80˚C.
3. Prepare working stock aliquots (0.07 ng/μl) by diluting 2 μl of immortal library in 28 μl nuclease-free water.

#### Prepare primers
+ Prepare forward and reverse primers working stocks(10µM) by diluting the 100µM stock in nuclease-free water.

#### Prepare water saturated diethyl ether.
+ Mix 100 ml of diethyl ether with 100 ml of nuclease-free water in a glass bottle by shaking for 30s. Allow the phases to settle before use (the top phase is diethyl ether).
+ The water saturated solvent can be stored at RT for 3 months.

## Procedure

### Prepare ePCR mix.

1. Prepare the oil-surfactant mixture by througly mixing the following components in a 50ml conical tube.

```{r, echo=FALSE}
oil_phase<-data.frame(Component= c("Span 80","Tween 80", "Triton X-100", "Mineral oil"), Amount=c("2.25ml", "200µl", "25µl", "to 50ml"),Final_conc=c("4.5% (vol/vol)","0.4% (vol/vol)","0.05% (vol/vol)",""))
colnames(oil_phase)<- c("Component","Amount", "Final concentration")
oil_phase %>%
  kable(escape=FALSE) %>%
  kable_styling(bootstrap_options = c("bordered","hover","striped", full_width = F, position="float-right"))
  
```

2. Transfer 400µl of the oil-surfactant mixure to a CryoTube vial, and add a 3x8 mm stir bar(pre-cleaned)
3. Begin stiring the mixture at 1,000 r.p.m. on the magnetic stirrer.
4. Prepare the aqueous phase for the emulsion by mixing the following:

```{r, echo=F}
aqua_phase<- data.frame(Component=c("10X PCR buffer",
                                    "BSA (100 mg/ml stock)",
                                    "Forward primer (10µM stock)",
                                    "Reverse primer (10µM stock)",
                                    "dNTPs (5mM stock)",
                                    "DNA polymerase",
                                    "Template DNA",
                                    "Nuclease-free water"),
                        Amount=c("26µl",
                                 "26µl",
                                 "7.8µl",
                                 "7.8µl",
                                 "10.4µl",
                                 "5.2µl",
                                 "≤10^9 molecules (1.66 fmol)",
                                 "to 260µl"))
aqua_phase %>%
  kable(escape=FALSE) %>%
  kable_styling(bootstrap_options = c("bordered","hover","striped", full_width = F, position="float-right"))
```
5. Add 200µl of the aqueous phase to the oil-surfactant mixure in dropwise manner over a period of 1.5 min.
    + After the addition is complete, continue stirring for another 5 minutes.
    + Keep the aqueous phase leftovers as the without emulsion contol.
6. Pipet the emulsion into the wells of a PCR plate as 10 aliquots of 50 µl.
    + Pipet 50 µl of the aqueos phase left over to a well as nonemulsified control.
    + __Overlay the emulsified and nonemulsified reactions with mineral oil.
### PCR

7. Subject the PCR plate to the following program.

| Cycle number | Denaturation | Annealing | Polymerization |
|--------------|--------------|-----------|----------------|
| 1            | 2 min at 95˚C |           |                |
| 2-26        | 30 sec at 95˚C | 30 sec at A^1^| B^1^ min at 72˚C|
| 27          |               |           |  10 min at 72 ˚C|

^1^ _A_ = _C_- 5˚C where _C_ is the Tm of the primer with lowest Tm. _B_ is equal to length of amplification in kbp.

8. Place the nonemulsified reaction aside.

--> __Pause Point:__ You can keep both the w and w/o emulsion reaction at 4˚C overnight!
_not sure about this!!_
### 




[^1]: https://www.ncbi.nlm.nih.gov/pubmed/16791213
[^2]: Version 1.4