---
title: "Preparing and labelling probe from a complex oligo library"
author: "Omid Gholamalamdari"
date: 'Updated 8/9/2018'
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo =F}
library(knitr)
library(kableExtra)
```

---

## Introduction

This protocol employs emulsion PCR (ePCR) to enrich sub libraries from a complex oligo library. I have prepared this protocol based on [ _"Amplification of complex gene libraries by emulsion PCR"_ ](https://www.ncbi.nlm.nih.gov/pubmed/16791213) _R. William, et.al, Nature Methods_[^1] and _"Preparing Single-Stranded Labelled Probes from MYtags Immortal Lirary"_[^2], Arbor Biosciences.
Labelling protocol is based terminal transferase protocol from Roche. 

### Materials
+ Oligo nucleotide library
+ Enrichment primer sets
+ Nuclease-free water
+ Span 80 (Sigma S6760)
+ Tween 80 (Sigma P5188)
+ Triton X-100 (Sigma T8787)
+ Minral Oil (Sigma M5904)
+ dNTP, 10mM (NEB N047S)
+ Bovine Serum Albumin (BSA) (Sigma B6917)
+ _Pfu_ Turbo DNA Polymerase
+ diethyl ether 


### Equipments
+ BRAND magnetic stirring bar, PTFE, cylindrical with pivot ring (Sigma Z329061). pre-cleaned (See ["Cleaning stir bars"](/cleaning_stir_bars.html)).
+ Magnetic stirrer with speed control (Burrell Scientific)
+ PCR cleanup kit

-----

## Before you begin

### Prepare library
1. Prepare a stock solution of your library. 
2. Resuspend at 1 ng/$\mu$l in 10 mM Tris-HCl pH 7.5 or nuclease-free water. 
    + Keep the stock at -80&deg; C.
3. Prepare working stock aliquots (0.08 ng/$\mu$l) by adding 2 $\mu$l of immortal library to 23$\mu l$ of nuclease-free water (total volume:25$\mu l$).

### Prepare primers
Prepare forward and reverse primers working stocks(10$\mu$M) by diluting the 100$\mu$M stock in nuclease-free water.

### Prepare water saturated diethyl ether.
Mix 100 ml of diethyl ether with 100 ml of nuclease-free water in a glass bottle by shaking for 30s. Allow the phases to settle before use (the top phase is diethyl ether). The water saturated solvent can be stored at RT for 3 months.

## Procedure
### Prepare ePCR mix.
1. Prepare the oil-surfactant mixture by througly mixing the following components in a 50ml conical tube.

```{r, eval=T,echo=FALSE}
oil_phase<-data.frame(Component= c("Span 80","Tween 80", "Triton X-100", "Mineral oil"), Amount=c("2.25ml", "200µl", "25µl", "to 50ml"),Final_conc=c("4.5% (vol/vol)","0.4% (vol/vol)","0.05% (vol/vol)",""))
colnames(oil_phase)<- c("Component","Amount", "Final concentration")
oil_phase %>%
  kable(booktab=T) %>%
  kable_styling(bootstrap_options = c("bordered","hover","striped", full_width = F, position="float-right"))
  
```

2. Transfer 400$\mu$l of the oil-surfactant mixure to a CryoTube vial, and add a 3x8 mm stir bar(pre-cleaned)
3. Begin stiring the mixture at 1,000 r.p.m. on the magnetic stirrer.
4. Prepare the aqueous phase for the emulsion by mixing the following:

| Component | Amount |
| ------| ---- |
|10X Cloned _Pfu_ buffer | 26$\mu l$ |
| BSA (100 mg/ml stock) | 26$\mu l$ |
| Forward primer (10µM stock) | 7.8$\mu l$ |
| Reverse primer (10µM stock) | 7.8$\mu l$ |
| dNTPs (10mM stock) | 5.2$\mu l$ |
| _Pfu_ Turbo DNA Polymerase | 5.2$\mu l$  |
| Template DNA | $\le 10^{9} molecules$ (1.66 fmol)  |
| Nuclease-free water | to 260 $\mu l$ |

5. Add 200$\mu$l of the aqueous phase to the oil-surfactant mixure in dropwise manner over a period of 1.5 min.
    + After the addition is complete, continue stirring for another 5 minutes.
    + Keep the aqueous phase leftovers as the without emulsion contol.
    
    $\blacktriangle$ __CRITICAL STEP__
6. Pipet the emulsion into the wells of a PCR plate as 10 aliquots of 50 $\mu$l.
    + Pipet 50 $\mu$l of the aqueos phase left over to a well as nonemulsified control.
    + Overlay the emulsified and nonemulsified reactions with mineral oil.
    
### PCR
7. Subject the PCR plate to the following program.

| Cycle number | Denaturation | Annealing | Polymerization |
|--------------|--------------|-----------|----------------|
| 1            | 2 min at 95&deg;C |           |                |
| 2-26        | 30 sec at 95&deg;C | 30 sec at A^1^| B^1^ min at 72&deg;C|
| 27          |               |           |  10 min at 72&deg;C|
_^1^A = C-5&deg;C where C is the Tm of the primer with lowest Tm. B is equal to length of amplification in kbp._
$\newline$

8. Place the nonemulsified reaction aside.

    $\blacksquare$ __PAUSE POINT__ You can keep both the w and w/o emulsion reaction at 4&deg;C overnight!

9. Pool the ePCR reactions in a 1.7 $ml$ eppendorf tube.
    + Centrifuge at 13,000$g$ for 5 minutes.
    + Discard the upper (oil) phase.

10. To remove the remaining oil from the emulsion, perform the folowing extraction twice:
    + Add 1 $ml$ of water-saturated diethyl ether, vortex the tube and let it settle.
    + Discard the upper (solvent) phase. 

    $\blacktriangle$ __CRITICAL STEP__

11. Remove the residual solvent from the broken emulsion by centrifuging under vaccum for 5 min at 25&deg;C. 
12. Clean the emulsified and nonemulsified PCR products using a PCR clean-up kit.
13. Analyze Aliquotes of the emulsified and nonemulsified PCR products by electrophoresis.
    + 2% agarose gel in 1X TBE.




[^1]: https://www.ncbi.nlm.nih.gov/pubmed/16791213
[^2]: Version 1.4